### 2017-07-01 UOW_thesis_template_v2-0/thesis_template/appendices/appendix1.tex
Committing core thesis files

\chapter{Derivation of useful lemmas from the preliminaries}
\chaptermark{Useful lemmas -- general}

\section{Proof of Lemma~\ref{lem:2}}
To prove the result, the Mellin transform will need to be evaluated directly. Using the definition of the Mellin transform and \eqref{eqn:Nprime}, the left-hand side becomes
	\begin{align*}
		\M \left\{ N'(a\log x + b) \right\} &= \int_0^\infty x^{\xi-1}N'(a\log x + b) \, \d x = \fr{1}{\sqrt{2\pi}}\int_0^\infty x^{\xi-1}e^{-(a\log x + b)^2/2}\, \d x.
	\end{align*}
Using $\omega = a\log x + b$, the integral becomes
	\begin{equation*}
		\M \left\{ N'(a\log x + b) \right\} = \fr{e^{-b\xi/a}}{a\sqrt{2\pi}}\int_{-\infty}^\infty e^{-\omega^2/2 + \omega\xi/a} \, \d\omega.
	\end{equation*}
Upon completing the square inside the exponential with respect to $\omega$, this simplifies to
	\begin{equation*}
		\M \left\{ N'(a\log x + b) \right\} =  \fr{e^{-b\xi/a}e^{\xi^2/(2a^2)}}{a\sqrt{2\pi}}\int_{-\infty}^\infty e^{-(\omega - \xi/a)^2/2} \, \d\omega =  \fr{e^{-b\xi/a}e^{\xi^2/(2a^2)}}{a},
	\end{equation*}
where we used the standard identity $\int_{-\infty}^\infty e^{-(\omega - y)^2/2} \, \d\omega = \sqrt{2\pi}$ for $y \in \mathbb{R}$. This completes the proof.

\section{Proof of Lemma~\ref{lem:1}}
To begin, we let $I$ equal
	\begin{equation*}
		I = \int_0^\infty \fr{1}{z}N'(a_1\log z + b_1)N'\left(a_2\log\left( \fr{1}{z} \right) + b_2 \right) \, \d z.
	\end{equation*}
Setting $\rho = a_2\log\left( 1/z \right) + b_2$, the integral becomes
	\begin{align*}
		I &= \fr{1}{a_2}\int_{-\infty}^\infty N'\left( \fr{a_1b_2}{a_2} + b_1 - \fr{a_1}{a_2}\rho \right)N'(\rho) \, \d \rho =  \fr{1}{a_2}\int_{-\infty}^\infty N'\left(\gamma - \fr{a_1}{a_2}\rho \right)N'(\rho) \, \d \rho,
	\end{align*}
where we write $\gamma = a_1b_2/a_2 + b_1$. Using \eqref{eqn:Nprime} to replace the $N'$ terms, we get
	\begin{align*}
		I &= \fr{1}{2\pi a_2}\int_{-\infty}^\infty e^{-(\left(\gamma- a_1\rho/a_2\right)^2 + \rho^2)/2} \, \d \rho \\
		&= \fr{e^{-\gamma^2/2}e^{(a_1\gamma)^2/(2(a_1^2+a_2^2))}}{2\pi a_2}\int_{-\infty}^\infty e^{-\left( \left(\left(a_1^2+a_2^2\right)/a_2^2\right)\left( \rho - (a_1a_2\gamma)/(a_1^2+a_2^2) \right)^2 \right)/2} \, \d \rho,
	\end{align*}
Now setting
	$$
		\omega = \sqrt{\fr{a_1^2+a_2^2}{a_2^2}}\left( u - \fr{a_1a_2\gamma}{a_1^2+a_2^2} \right),
	$$
we obtain
	$$
		I = \fr{e^{-\gamma^2/2}e^{(a_1\gamma)^2/(2(a_1^2+a_2^2))}}{2\pi \sqrt{a_1^2+a_2^2}}\int_{-\infty}^\infty e^{-\omega^2/2} \, \d\omega.
	$$
To finish off, we use \eqref{eqn:Nprime} to replace the exponential term with $N'$, integrate, and arrive at
	\begin{align*}
		I &= \fr{1}{\sqrt{a_1^2 + a_2^2}}N'\left( \fr{\gamma a_2}{\sqrt{a_1^2+a_2^2}} \right) =\fr{1}{\sqrt{a_1^2 + a_2^2}}N'\left( \fr{a_1b_2 + a_2b_1}{\sqrt{a_1^2+a_2^2}} \right),
	\end{align*}
where we substituted the expression for $\gamma$ back in. This completes the proof.

\section{Proof of Lemma \ref{lem:2b}}
The first step is to let $I$ be equal to the integral. Setting $\rho = a_2\log\left( 1/z \right) + b_2$ gives
	\begin{align*}
		I &= \fr{1}{a_2}\int_{-\infty}^{a_2\log(1/c)+b_2} N'\left( \fr{a_1b_2}{a_2} + b_1 - \fr{a_1}{a_2}\rho \right)N'(\rho) \, \d \rho \\
		&=  \fr{1}{a_2}\int_{-\infty}^{a_2\log(1/c)+b_2} N'\left(\gamma - \fr{a_1}{a_2}\rho \right)N'(\rho) \, \d \rho,
	\end{align*}
when we have assigned $\gamma = a_1b_2/a_2 + b_1$. The $N'$ terms now become replaced with their exponential expressions via~\eqref{eqn:Nprime}, which changes $I$ to be
	\begin{align*}
		I &= \fr{1}{2\pi a_2}\int_{-\infty}^{a_2\log(1/c)+b_2} e^{-(\left(\gamma- a_1\rho/a_2\right)^2 + \rho^2)/2} \, \d \rho \\
		&= \fr{e^{-\gamma^2/2}e^{(a_1\gamma)^2/(2(a_1^2+a_2^2))}}{2\pi a_2}\int_{-\infty}^{a_2\log(1/c)+b_2} e^{-\left( \left(\left(a_1^2+a_2^2\right)/a_2^2\right)\left( \rho - (a_1a_2\gamma)/(a_1^2+a_2^2) \right)^2 \right)/2} \, \d \rho,
	\end{align*}
Now we make another substutition
	$$
		\omega = \sqrt{\fr{a_1^2+a_2^2}{a_2^2}}\left( u - \fr{a_1a_2\gamma}{a_1^2+a_2^2} \right),
	$$
and this gives
	$$
		I = \fr{e^{-\gamma^2/2}e^{(a_1\gamma)^2/(2(a_1^2+a_2^2))}}{2\pi \sqrt{a_1^2+a_2^2}}\int_{-\infty}^d e^{-\omega^2/2} \, \d\omega,
	$$
with
	$$
		d =  \sqrt{a_1^2+a_2^2}\left( \log(1/c) + \fr{b_2}{a_2} - \fr{a_1\gamma}{a_1^2+a_2^2} \right).
	$$
Finally, we use \eqref{eqn:Nprime} to replace the exponential term with $N'$, integrate, and arrive at
	\begin{align*}
		I &= \fr{1}{\sqrt{a_1^2 + a_2^2}}N'\left( \fr{\gamma a_2}{\sqrt{a_1^2+a_2^2}} \right)N(d) =\fr{1}{\sqrt{a_1^2 + a_2^2}}N'\left( \fr{a_1b_2 + a_2b_1}{\sqrt{a_1^2+a_2^2}} \right)N(d),
	\end{align*}
Thus, the proof is concluded.

\section{Proof of Lemma \ref{lem:2bb}}
The proof for this lemma is nearly identical to the proof for lemma~\ref{lem:2b}. Therefore we will only provide a sketch of the proof. We once first let $I$ be equal to the integral. Making the substitution $\rho = a_2\log(1/z) + b_2$ gives
	\begin{equation*}
		I =  \fr{1}{a_2}\int_{a_2\log(1/c)+b_2}^\infty N'\left(\gamma - \fr{a_1}{a_2}\rho \right)N'(\rho) \, \d \rho,
	\end{equation*}	
where the quantity $\gamma$ is equal to $a_1b_2/a_2 + b_1$. Using~\eqref{eqn:Nprime} to replace the $N'$ terms with their expontential counterparts and then using the substitution
	$$
		\omega = \sqrt{\fr{a_1^2+a_2^2}{a_2^2}}\left( u - \fr{a_1a_2\gamma}{a_1^2+a_2^2} \right),
	$$
it follows that
	$$
		I = \fr{e^{-\gamma^2/2}e^{(a_1\gamma)^2/(2(a_1^2+a_2^2))}}{2\pi \sqrt{a_1^2+a_2^2}}\int_d^\infty e^{-\omega^2/2} \, \d\omega,
	$$
with
	$$
		d =  \sqrt{a_1^2+a_2^2}\left( \log(1/c) + \fr{b_2}{a_2} - \fr{a_1\gamma}{a_1^2+a_2^2} \right).
	$$
The final step now is to use~\eqref{eqn:Nprime} to allow $N'$ in place of the exponential term, integrate, and use the properties of the normal CDF $N$ to get
	\begin{align*}
		I  &=\fr{1}{\sqrt{a_1^2 + a_2^2}}N'\left( \fr{a_1b_2 + a_2b_1}{\sqrt{a_1^2+a_2^2}} \right)N(-d),
	\end{align*}

\section{Proof of Lemma \ref{lem:2a}}
The proof of the first expression is simple. We denote the left-hand side integral to be $I$. By setting $u = a_1\log(1/y) + b_1$, we get
	$$
		I = \fr{1}{a_1}\int_{a_1\log(1/b)+b_1}^{a_1\log(1/a)+b_1} N'(u) \, \d u = \fr{1}{a_1}\bigg( N(a_1\log(1/a)+b_1) - N(a_1\log(1/b)+b_1) \bigg).
	$$
For the second expression, we perform the exact same step as we did for the first expression to begin with. We first obtain
	$$
		I = \fr{e^{b_1/a_1}}{a_1}\int_{a_1\log(1/b)+b_1}^{a_1\log(1/a)+b_1} N'(u)e^{-u/a_1} \, \d u = \fr{e^{b_1/a_1}}{a_1\sqrt{2\pi}}\int_{a_1\log(1/b)+b_1}^{a_1\log(1/a)+b_1} e^{-u^2/2-u/a_1} \, \d u,
	$$
where we used~\eqref{eqn:Nprime} to convert $N'$ to its integral form. Looking at the power of the exponential term, we complete the square and arrive at
	\begin{align*}
		I &= \fr{e^{b_1/a_1+1/(2a_1^2)}}{a_1\sqrt{2\pi}}\int_{a_1\log(1/b)+b_1}^{a_1\log(1/a)+b_1}e^{-(u+1/a_1)^2/2} \, \d u \\
		&= \fr{e^{b_1/a_1+1/(2a_1^2)}}{a_1}\int_{a_1\log(1/b)+b_1}^{a_1\log(1/a)+b_1}N'\left( u + 1/a_1 \right) \, \d u \\
		&= \fr{e^{b_1/a_1 + 1/(2a_1^2)}}{a_1}\bigg( N(a_1\log(1/a)+b_1 + 1/a_1) - N(a_1\log(1/b) +b_1 + 1/a_1) \bigg).
	\end{align*}
This equals the right-hand side of the second expression and concludes the proof.