### 2017-07-01 UOW_thesis_template_v2-0/thesis_template/frontmatter/acknowledgments.tex

\chapter{Acknowledgements}

First and foremost, I would like to extend my sincerest gratitude to my primary supervisor, Marianito Rodrigo. You managed to draw out a potential in me that I never thought was there. Through a mixture of you applying pressure and maintaining mathematical rigor, we eventually pulled through. Thank you for being patient throughout the last three years.

Secondly, I would like to thank Joanna Goard (my co-supervisor) for always having time to read my drafts and providing great mathematical insight whenever I inadvertently left everything to the last minute. Like Marianito, thank you for also exercising patience in my times of carelessness.

To all the administrative staff, fellow PhD students, and budding academics who always stopped to have a chat with me, I thank you all for your spontaneity in our haphazard discourses.

All that I am today would not have been possible without the unconditional love from my parents, Huijun Li (dad) and Shaohua Zhou (mum). Dad, you were extremely hesitant about me deciding to do a PhD as you went through it yourself and understand the seething stress behind it all. But you always had a smile on your face and managed to cheer me up whenever I was feeling extremely dissatisfied with everything involving my PhD. Thank you for always reassuring me that everything was going to be okay. Mum, you too were also cautious about me pursuing a PhD as you’ve seen how much dad had to go through all those years ago. Like dad however, you always kept encouraging me to finish this chapter of my life and to make sure I never had any regrets with any of my decisions. Thanks for always telling me what I needed to hear, because what I wanted to hear would not have been beneficial at times.

I definitely need to extend a big appreciation and express my utter disbelief at the support I have received from my “stream fam” over at twitch.tv/MikamiHero. You all gave me an outlet nearly every single night for the last two years to just hang out, play some video games, share a few laughs and most of all, get my mind away from my PhD. I truly meant it when I said that if I had never started streaming on 30 December 2014, I would’ve most certainly quit the PhD program. You guys have provided me something far more special than just being viewers. 
\#mikamiLove

No amount of words can vocalize how much I value the true love and unwavering support from my long-term girlfriend, Marjorie McDonald. You too were doing a PhD and thus we shared similar parallels in terms of stress and perpetual worry. Yet we always managed to set aside time for each other, be it having dinner, watching a movie, going for a swim. The list is endless. We were also always there for one another in very dire circumstances. I could not have done this without you, my love. Thank you for everything.

Lastly, I would like to thank God. Nothing I ever say will be worthy of His greatness, but without Him, I am literally nothing. Thank you for always forgiving me and providing the faith I required to persevere whenever blunders happened. To quote Galatians 6:9, “Let us not become weary in doing good, for at the proper time we will reap a harvest if we do not give up.”